A Container-Based Cloud-Native Architecture for the Reproducible Execution of Multi-population Optimization Algorithms

### 2020-10-17 draft.tex
Update draft.tex

\begin{table}[h!tbp]
  \caption{Speed-up by worker and dimension, taking one worker as the
    baseline. Speed-up is obtained by dividing the evaluation ratio of the baseline by the 
    ratio of the increased number of workers.}
% Clarify here what you are measuring. Number of evaluations for
% whatever divided by whatever - JJ
% Done - Mario
  \label{tab:speedup-table}
  \vspace{0.25cm}
  \centering

@@

%
As expected, within the same number of populations, there is a linear speed-up
when we add more workers. The speed-up is marginal in those cases where there 
are fewer populations than workers.  The reason for this is that at any given time, 
there are at least two populations in the message queue, leaving some of the workers 
waiting for populations to be available in the queue.

@@

Table~\ref{tab:speedup:test} shows the speed-up obtained taking as the baseline
the previous configuration, for example, when increasing from four to eight
workers, and the p-value for the Wilcoxon Signed-Rank test. There was an
increment in all cases. Again, the best effect is achieved when the number of
populations is higher than the number of workers. Configurations following this
basic rule doubled the speed or were near that. Using five populations, we
obtained the best and worst speed-up depending on the number of workers (shown in
underline and bold, respectively).

@@

In the previous section, we have shown that as we increase the number of
populations and workers, we increase the ratio of the number of evaluations per
second in the same proportion. However, it is vital to notice that we can also
affect the number of evaluations we need to execute to find a solution. This
reduction means that the multi-population based algorithm can perform better
when we increase the number of populations, or in some cases, even workers, as the experiments
show. If we increase the evaluations per second and, at the same time, decrease
the number of evaluations needed, we can have a superlinear speed-up of the
algorithm, if we measure the time required to complete an instance.

@@

First, we will see how the configuration we selected can reduce the number of
evaluations needed in an experiment. Figure \ref{fig:evals_instance} shows the
median (30 runs) of the number of evaluations needed to complete an instance of
the function $f_4$. For every selection in the number of populations, we could
expect to have about the same number of FEs needed regardless of the number of
workers. However, we can see that for some configurations, i.e., 10D and ten
populations, eight workers will need less FEs then one (p = 0.012). Also, for
20D and twenty populations, there are differences between one and four workers
(p = 0.08) and between four and sixteen (p = 0.01). These results indicate that
there are benefits in the potential change of order in the message queue
resulting from the parallelization of work.

@@

The reduction in the number of FEs has an impact when we measure the speed up
concerning the time needed to complete an experiment. In this case, we have
measured the median of the seconds needed to complete an instance of the
problem. Figure \ref{fig:spworker:time} shows in each box-plot the median (30
runs) of the execution time achieved by 2, 4, 8, and 16 workers, to finish an
instance of the function $f_4$, for dimensions 10 and 20, when using 5, 10, and
20 population sizes. We present each dimension in a row having a log scale on
the y-axis, to compensate for the dimension increments and scalability.

@@

In Table~\ref{tab:time} we can see that in some cases we have a supralinear
speed-up, we highlighted these in boldface and indicated the speed-up value in
parenthesis. The best overall times are underlined. As expected, increasing the
number of workers for a fixed dimension and populations reduced the time
required to find a solution, as long as the population number was higher than
the number of workers. The values that had no statistical difference (in the
Wilcoxon test) when increasing the number of workers have an asterisk.

>>>>

\begin{table}[h!tbp]
  \caption{Speed-up by worker and dimension, taking one worker as the
    baseline. Speed-up is obtained by dividing the baseline's evaluation ratio by the 
    ratio of the increased number of workers.}
% Clarify here what you are measuring. Number of evaluations for
% whatever divided by whatever - JJ
% Done - Mario
  \label{tab:speedup-table}
  \vspace{0.25cm}
  \centering

@@

%
As expected, there is a linear speed-up when we add more workers 
within the same number of populations. The speed-up is marginal in those cases where there 
are fewer populations than workers. This is because, at any given time, 
there are at least two populations in the message queue, leaving some of the workers 
waiting for populations to be available in the queue.

@@

Table~\ref{tab:speedup:test} shows the speed-up obtained taking as the baseline
the previous configuration, such as increasing from four to eight
workers, and the p-value for the Wilcoxon Signed-Rank test. There was an
increment in all cases. Again, the best effect is achieved when the number of
populations is higher than the number of workers. Configurations following this
basic rule doubled the speed or where near that. Using five populations, we
obtained the best and worst speed-up depending on the number of workers (shown in
underline and bold, respectively).

@@

In the previous section, we have shown that as we increase the number of
populations and workers, we increase the number of evaluations per
second in the same proportion. However, it is vital to notice that we can also
affect the number of evaluations we need to execute to find a solution. This
reduction means that the multi-population based algorithm can perform better
when we increase the number of populations, or in some cases, even workers, as the experiments
show. If we increase the evaluations per second and, at the same time, decrease
the number of evaluations needed, we can have a superlinear speed-up of the
algorithm, if we measure the time required to complete an instance.

@@

First, we will see how the configuration we selected can reduce the number of
evaluations needed in an experiment. Figure \ref{fig:evals_instance} shows the
median (30 runs) of the number of evaluations needed to complete an instance of
the function $f_4$. For every selection in the number of populations, we could
expect to have about the same number of FEs needed regardless of the number of
workers. However, we can see that for some configurations, i.e., 10D and ten
populations, eight workers will need fewer FEs than one (p = 0.012). Also, for
20D and twenty populations, there are differences between one and four workers
(p = 0.08) and between four and sixteen (p = 0.01). These results indicate that
there are benefits in the potential change of order in the message queue
resulting from work parallelization.

@@

The reduction in the number of FEs impacts when we measure the speed up
concerning the time needed to complete an experiment. In this case, we have
measured the median of the seconds needed to complete an instance of the
problem. Figure \ref{fig:spworker:time} shows in each boxplot the median (30
runs) of the execution time achieved by 2, 4, 8, and 16 workers, to finish an
instance of the function $f_4$, for dimensions 10 and 20, when using 5, 10, and
20 population sizes. We present each dimension in a row having a log scale on
the y-axis to compensate for the dimension increments and scalability.

@@

In Table~\ref{tab:time} we can see that we have a supralinear
speed-up in some cases; we highlighted these in boldface and indicated the speed-up value in
parenthesis. The best overall times are underlined. As expected, increasing the
number of workers for a fixed dimension and populations reduced the time
required to find a solution, as long as the population number was higher than
the number of workers. The values that had no statistical difference (in the
Wilcoxon test) when increasing the number of workers have an asterisk.
